Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-29 bbbq_article.tex
Process feedback Geert, thanks @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/196

Using a alpha value of $0.05$, as is common practice for an uninformed
expectation, we conclude that there are significantly less SNPs
in TMHs as expected by chance.
See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
when randomly selecting SNPs,
within TMPs that are 50\% TMH (or correcting for this percentage),
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}. [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}. Ik denk dat je deze hieronder kan samenvoegen]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{TMH-derived epitopes are predicted to be presentated in vivo}

@@

Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown (\cite{bianchi2017}). Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved. [GEERT: beetje kort. Suggesties? Experimental approaches? Hoeft niet super origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. Frans kan vast helpen met mooie zin.]

@@

An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
However,defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
and peptide fusion (\cite{delong2016pathogenic}) are
two processes that both result in peptide fragments
that did not originate from the mature proteins. [GEERT: mooi argument, maar moet wel kort uitgewerkt worden. Uitleggen wat DRiPs zijn en wat peptide fusion is, en hoe het kan leiden tot false positives]

@@

\subsection{TMH-derived epitopes are presented more often than expected by chance}

@@

\subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}

@@

In general, one would hope that evolutionary selection results in
an immune system that as most attentive for protein regions that are
essential for the survival, proliferation and/or virulence or pathogenic microbes, as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that change its 
virulence (\cite{velazquez2020positive}).
These regions, however, only account for a small part of a pathogen's proteome.
Additionally, these essential parts might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible to provide innate immune responses against such essential protein regions, 
as hinted upon by a study on influenza (\cite{han2019individual}),
where it was found that the selection pressure
exerted by the immune system was a non profounding factor.

@@

TMHs, on the other hand, also have strict functional constraints as they need to span through hydrophobic lipid bilayers. Moreover, TMHs are abundant and make up ... of a human, bacterial or viral proteome.
Proteomes thus contain many TMHs,
and it might therefore be beneficial for host organisms
to tailor their immune systems towards immune responses against foreign TMHs.
This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focusing on TMHs only, if a signal of selection
can be detected.\frans{ik snap deze redenatie niet?}[GEERT: ik ook niet, vind beetje onnodig. Volstaat te zeggen dat TMHs niet de enige hydrofobe regio's zijn, en ook belang voor intra-protein en protein-protein interacties en catalytische sites. Daarnaast in deze paragraaf iets zeggen over single vs multi-TMH eiwitten]
We speculate that the human immune system is more attentive 
towards TMHs, as it is a general structure, that is evolutionarily conserved.

[GVDB: 
  Also compare your findings with previous efforts 
  to look at conservation of TMHs
]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Supplementary materials}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Differences with Bianchi et al., 2017}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A part of this study does the same analysis as Bianchi et al., 2017.mainly concern the use of different
software and a different definition of what an MHC binder is.

% \paragraph{IC50 prediction software}

The earlier study used \verb;epitope-prediction; a hand-crafted method [GEERT: what is hand-crafted? They didn't use a computer? Why relevant?], 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'mhcnuggetsr' R package was used in this study.

>>>>

Using a alpha value of $0.05$, as is common practice for an uninformed
expectation, we conclude that there are significantly less SNPs
in TMHs as expected by chance.
See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
when randomly selecting SNPs,
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}. [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}. Ik denk dat je deze hieronder kan samenvoegen]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived epitopes are predicted to be presentated in vivo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown (\cite{bianchi2017}). Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved.
\geert{
  beetje kort. Suggesties? Experimental approaches? Hoeft niet super
  origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. 
  Frans kan vast helpen met mooie zin
}

@@

An important caveat of the elution studies is that we assume that 
the found epitopes are derived from their completely 
translated predecessor proteins.
However, there are multiple cellular processes 
that may result in peptide fragments that did not originate 
from the mature proteins.
One such process is the production of defective ribosomal 
products (DRiPs) (\cite{yewdell1996defective}),
in which the ribosome terminates its translation prematurely,
intrudes a point mutation, 
or cause a frame shift, at low frequencies (\cite{yewdell2011drips}).
Another process that results in peptide fragments absent in the
proteome is peptide fusion, 
in which peptides form covalent bonds with other peptides,
as is for example found in mice (\cite{delong2016pathogenic}).
Due to these processes, we may incorrectly attribute an
epitope to a specific proteome locations, 
where in reality, it was created de novo.
This may lead to false positives, as we incorrectly deduce
epitopes to be derived from a TMH, instead of from one of
these processes. Both processes, however, are thought to
occur at low frequencies (\cite{yewdell2011drips,delong2016pathogenic}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived epitopes are presented more often than expected by chance}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In general, one would hope that evolutionary selection results in
an immune system that as most attentive for protein regions that are
essential for the survival, proliferation and/or virulence or pathogenic microbes, as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that change its 
virulence (\cite{velazquez2020positive}).
These regions, however, only account for a small part of a pathogen's proteome.
Additionally, these essential parts might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible to provide innate immune responses against such essential protein regions, 
as hinted upon by a study on influenza (\cite{han2019individual}),
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

@@

Evolutionary selection of pathogens by a host's immune system,
however, is likelier to occur for proteomic patterns that are general,
over patterns that are rare.
Where essential catalytic sites in a pathogenic proteome
are rare, TMHs are common and thus a more feasible 
target for evolution to respond to.
We have found the signature of evolution when both factors,
that is, TMHs and catalytic sites are likely to co-occur,
which is in proteins that span the membrane at least twice.
This may be explained additionally that multi-spoanners
are not only likelier to have a catalytically active site,
also there are interactions between the TMHs.
For single-spanners, we found no evolutionary conservation.
We speculate that the human immune system is more attentive 
towards TMHs in multi-spanners, as these are evolutionarily conserved.

\geert{
  Also compare your findings with previous efforts 
  to look at conservation of TMHs
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Supplementary materials}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Differences with Bianchi et al., 2017}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A part of this study does the same analysis as Bianchi et al., 2017.
mainly concern the use of different
software and a different definition of what an MHC binder is.
